\documentclass[12pt]{article}

\usepackage{fullpage,url,amsmath,amsthm,amssymb,epsfig,color,xspace,sigsam,mathrsfs}
\usepackage[pdftitle={abstract}]{hyperref}

\title{Sparse FGLM using the block Wiedemann algorithm}
\author{Seung Gyu Hyun, Vincent Neiger, Hamid Rahkooy, \' Eric Schost}

\titlehead{Block Sparse FGLM Algorithm}
\articlehead{ISSAC 2017 poster abstract}
\authorhead{Seung Gyu Hyun, Vincent Neiger, Hamid Rahkooy, \' Eric Schost}

\begin{document}

\maketitle

\noindent{\bf Overview.}
Computing the Gr\"obner basis of an ideal with respect to a term
ordering is an essential step in solving systems of polynomials; in
what follows, we restrict our attention to systems with finitely many
solutions. Certain term orderings, such as the degree reverse
lexicographical ordering (\emph{degrevlex}), make the computation of the
Gr\"obner basis faster, while other orderings, such as the
lexicographical ordering (\emph{lex}), make it easier to find the coordinates
of the solutions. Thus, one typically first computes a Gr\"obner basis
for the degrevlex ordering, and then converts it to either a lex Gr\"obner
basis or a related representation, such as Rouillier's Rational
Univariate Representation~\cite{Rouillier99}.

Consider a zero-dimensional ideal $I \subset \mathbb{K}[x_1, \dots,
  x_n]$, given by means of a monomial basis $\mathscr{B}$ of
$Q=\mathbb{K}[x_1, \dots, x_n]/I$ together with the multiplication
matrices $T_1,\dots, T_n \in \mathbb{K}^{D \times D}$ of
$x_1,\dots,x_n$ in $Q$ in $\mathscr{B}$, with
$D=\dim_\mathbb{K}(Q)$. In all that follows, we assume that $x_n$ {\em
  separates} the points of $V(I)$. However, we assume neither that $I$
is radical nor that it is in {\em shape position}; we recall that the latter means that the
Gr\"obner basis of $I$ for the lex ordering with $x_1 > \cdots > x_n$ has the form
$(x_1-G_1(x_n),\dots,x_{n-1}-G_{n-1}(x_n),G_n(x_n))$. Note that our
assumption can be ensured by a generic change of coordinates, while
the shape position one may not hold for any choice of
coordinates. Under our assumption, the {\em radical} of $I$ is in
shape position, with lex Gr\"obner basis
$R=(x_1-R_1(x_n),\dots,x_{n-1}-R_{n-1}(x_n),R_n(x_n))$.

The FGLM algorithm~\cite{FaGiLaMo93} computes the Gr\"obner basis of
$I$ for the lex ordering with a runtime cubic in the dimension
$D$. Although recent work has reduced the runtime exponent from $3$ to
the exponent of matrix multiplication
$\omega$~\cite{FaGaHuRe13,Neiger16}, linear algebra techniques based
on duality often turn out to be more efficient. The basic idea behind
these techniques is to compute values of a
linear form $\ell: Q \to \mathbb{K}$ at well-chosen elements of $Q$,
or possibly of several such forms,
and to deduce our output by solving some well-structured linear system
(typically, a Hankel system).

Several forms of this idea exist. The Rational Univariate
Representation algorithm takes for $\ell$ the trace ${\rm tr}: Q \to
\mathbb{K}$ and computes values of the form ${\rm tr}(x_n^i)$ and
${\rm tr}(x_j x_n^i)$; from there, the output of the algorithm is
equivalent to the data of $R$, together with the multiplicities of all
points. Values such as ${\rm tr}(x_n^i)$ can be computed as $v T_n^i
e_n$, where $v$ is the row vector of the traces of the monomial basis
$\mathscr{B}$, and $e_n$ is the column vector of coordinates of $x_n$ on
$\mathscr{B}$. Computing the values of the trace on $\mathscr{B}$ is
however costly, so several algorithms use random linear forms
instead. In~\cite{BoSaSc03}, the authors show how we can recover the
Gr\"obner basis $R$ of $\sqrt{I}$ (together with the nil-indices of
all points) from the values $\ell(x_n^i)$ and $\ell(x_j x_n^i)$, for a
random $\ell$, whereas Faug\`ere and Mou show in~\cite{FaMo17} how to
compute the basis $G$ of $I$ using similar sets of values, when
$I$ is in shape position (in general, the algorithm falls back on
Berlekamp-Massey-Sakata techniques). Quite importantly, they also
show that exploiting the sparsity of the matrices $T_1,\dots,T_n$
is critical for the efficiency of these algorithms.

The computation of sequences such as $\ell(x_n^i)$ is difficult to
parallelize.  In this work, in the continuation of~\cite{BoSaSc03},
we use techniques inspired by Coppersmith's block Wiedemann
algorithm~\cite{Coppersmith93} to compute the lex Gr\"obner basis $R$ of
$\sqrt{I}$.

\medskip\noindent{\bf Computing $R_{n}$.} We choose an integer $m$, random
matrices $u \in \mathbb{K}^{m \times D}$, $v \in \mathbb{K}^{D \times
  m}$, and we consider the matrix sequence $s = (uT_n^iv)_{i \ge
  0}$. For generic matrices $u, v$, \cite[Theorem 2.12]{KaVi04}
shows that if $S \in \mathbb{K}[x]^{m\times m}$ is a minimal
generating polynomial of $(s_i)_{i \ge 0}$, its determinant is a
multiple of the minimal polynomial of $x_n$ in $Q$, and divides its
characteristic polynomial. This polynomial must then factor as
$\det(S) = \prod_{\alpha=(\alpha_1,\dots,\alpha_n) \in V(I)}
(x-\alpha_n)^{e_\alpha}$, for some positive integers $(e_\alpha)_\alpha$; in
particular, its squarefree part (written in the variable $x_n$) is the last polynomial $R_n$ in $R$.

\medskip\noindent{\bf Using $u$-resultant techniques.} We propose  to recover the 
whole Gr\"obner basis $R$ of $\sqrt{I}$ by computing a truncation of
a polynomial akin to the $u$-resultant; we show here this idea for the computation
of $R_{n-1}$.

Let $\lambda$ be a new variable and consider the matrix $T_\lambda=T_n
+ \lambda T_{n-1}$; this is the multiplication matrix by $x_n +
\lambda x_{n-1}$ in $Q \otimes_{\mathbb{K}}
\mathbb{K}(\lambda)$. Applying the above algorithm to the sequence
$s_\lambda=(uT_\lambda^iv)_{i \ge 0}$ gives us the polynomial
$M_\lambda=\prod_{\alpha=(\alpha_1,\dots,\alpha_n) \in V(I)}
(x-(\alpha_n+\lambda \alpha_{n-1}))$; it is well-known that the
polynomial $R_{n-1}$ can be deduced from the coefficients of
$\lambda^0$ and $\lambda^1$ in $M_\lambda$ in quasi-linear time.

Hence, it is enough to compute $M_\lambda \bmod \lambda^2$. For this,
we compute the sequence $s_\lambda \bmod \lambda^2=(s_i + i s^*_i \lambda)_{i \ge 0}$,
with $s^*_i= uT_n^{i-1} T_{n-1}v$, and we apply e.g.\ the matrix
Berlekamp Massey algorithm to this sequence, over the coefficient ring
$\mathbb{K}[\lambda]/(\lambda^2)$; in generic coordinates, we expect that
the calculation carries over as if we were over a field.

\medskip\noindent{\bf The case of radical ideals.} It is of course desirable to 
avoid computing minimal matrix polynomials over non-reduced rings, if
possible. For radical ideals, we propose an alternative solution that
avoids this issue (extending this idea to arbitrary $I$ is work in progress).

Suppose that we have computed the minimal generating polynomial $S$,
together with its determinant (which, in the radical case, coincides with
$R_n(x_n)$). We can then compute the numerator matrix $N$ of the
generating series $G=\sum_{i \ge 0} s_i/x^{i+1}$ by a single product
of polynomial matrices $N=S G$. We repeat this operation with the
generating series $\sum_{i \ge 0} s^*_i/x^{i+1}$, for $s^*_i$ as in
the previous paragraph, to obtain a numerator matrix $N^*$. The Smith
form of $S$, $D=A S B$, only has $R_n(x)$ as a non-trivial
invariant factor. Then, let $n$ and $n^*$ be the entries of indices
$(1,m)$ in respectively $AN$ and $AN^*$; we can verify that
$R_{n-1}=n^*/n \bmod R_n$.

As an example, we will run the algorithm on $I = \langle f_1, f_2, f_3 \rangle \subset \mathbb{F}_{9001}[x_1,x_2,x_3]$,
where
{\small
\begin{align*}
f_1 &= 3536x_3^2 +6536x_3x_2 + 3900x_2^2 + 3722x_3x_1 + 580x_2x_1 + 7635x_1^2 + 4203x_3 + 1386x_2 + 2491x_1 + 250\\
f_2 &= 3987x_3^2 + 3953x_3x_2 +6122x_2^2 +5115x_3x_1 +7660x_2x_1 +8669x_1^2 + 4098x_3 +7705x_2 + 2449x_1 + 1134\\
f_3 &= 8589x_3^2 + 1291x_3x_2 +5321x_2^2 + 765x_3x_1 +6052x_2x_1 +8178x_1^2 + 2764x_3 + 957x_2 +7079x_1 + 517.
\end{align*}
}%
We choose $m = 2$ and two random matrices
\begin{align*}
u &= \begin{bmatrix}
291&  22& 337& 924& 414& 666& 574& 707\\
57& 801& 513& 135& 447& 107& 942& 320
\end{bmatrix}\\
v &= \begin{bmatrix}
553&  81&  56& 261& 109& 890& 477&  53\\
725& 642& 905& 612& 952& 158& 235& 783
\end{bmatrix} ^ {t} .
\end{align*}
(In this specific example, we have restricted to random coefficients in $\{0,\ldots,999\}$ from the prime field $\mathbb{F}_{9001}$.)
The monomial basis of $Q$ is $\mathscr{B}=[x_1^3, x_1^2, x_2x_1, x_3x_1, x_1, x_2, x_3, 1]$, with $D = 8$. 
Computing the minimal generating polynomial of $s=(u T_3^i v)_{i\ge 0}$, we get
$$S= \begin{bmatrix}
8226 + 5622x + 7693x^2 + 5033x^3 + x^4&     2919 + 8706x + 3736x^2 + 4829x^3\\
7928 + 3675x + 7471x^2 + 3510x^3  &        3113 + 5615x + 2702x^2 + 3353x^3 + x^4
\end{bmatrix},$$
with determinant
$$R_3 =x^8 + 8386x^7 + 8262x^6 + 7301x^5 + 318x^4 + 4870x^3 + 715x^2 + 8568x + 8433.$$
We compute the numerator matrix $N$ of the generating series $\sum_{i \ge 0} s_i/x^{i+1}$
and multiply it by the change of basis of $A$ obtained from the Smith form computation of 
$S$; the $(1,2)$-entry in the result is
$$n = 5527x^7 + 2064x^6 + 4391x^5 + 1308x^4 + 6797x^3 + 1328x^2 + 6317x + 8700.$$
Computing the numerator matrix $N^*$ of the generating series $\sum_{i \ge 0} s'_i/x^{i+1}$,
with $s'_i = uT_3^{i-1} T_2 v$, and doing as above, we obtain
$$n^* = 8649x^7 + 3840x^6 + 4938x^5 + 1734x^4 + 1525x^3 + 2362x^2 + 1780x + 2397.$$
The polynomial $R_2(x_3)$ is then given by $n^*/n \bmod R_3$, written in the variable
$x_3$, and
is equal to 
$$R_2 = 2287x_3^7 + 8269x_3^6 + 8475x_3^5 + 6889x_3^4 + 4785x_3^3 + 3960x_3^2 + 3902x_3 + 4153.$$
Doing the same with the sequence $uT_3^{i-1} T_1 v$ gives us a polynomial $R_1(x_3)$,
which finally leads us to the lex Gr\"obner basis $(x_1-R_1(x_3), x_2-R_2(x_3),R_3(x_3))$
of $I$.

%% It is possible to verify the correctness of this algorithm algebraically, given that the Smith normal form of
%% S is as described. We know experimentally that this is the case, but we do not know how to prove this relationship
%% between the Smith normal form and the least common left multiple matrix $S$. Further work must also be done to
%% make the algorithm work for ideals that are not radical.

\bibliographystyle{plain}
\bibliography{abstract}


\end{document}

